\begin{changelog}
\change{2021-02-16}{Kevin Hammond}{FM (IOHK)}{Initial version. }
\change{2021-02-17}{Kevin Hammond}{FM (IOHK)}{Description of goals and submission process.}
\change{2021-02-19}{Kevin Hammond}{FM (IOHK)}{Groups involved.  Vote delegation.}
\change{2021-02-19}{Kevin Hammond}{FM (IOHK)}{Added various outline sections.  Voting, sidechains, short Priviledge comparison.}
\change{2021-02-22}{Kevin Hammond}{FM (IOHK)}{Added workflows, plus textual improvements.}
\change{2021-02-25}{Kevin Hammond}{FM (IOHK)}{Added transition process, updated diagrams, described endorsement and enactment, cleaned up text.}
\change{2021-03-01}{Kevin Hammond}{FM (IOHK)}{Reviewed and updated text.  Added missing diagram, section on transition.}
\change{2021-03-03}{Kevin Hammond}{FM (IOHK)}{Added appendix on user experience.}
\change{2021-03-05}{Kevin Hammond}{FM (IOHK)}{Revised and tidied.  Added address structure and other diagrams.  Checked against delegation design document.}
\change{2021-03-10}{Kevin Hammond}{FM (IOHK)}{Clarified automated endorsement.  Decision needs to be taken on whether endorsement is stake-based or block-based.}
\change{2021-03-12}{Kevin Hammond}{FM (IOHK)}{Added appendix on protocol parameters.  Updated text.}
\change{2021-03-12}{Kevin Hammond}{FM (IOHK)}{Incorporated initial feedback from Andre and Philipp.}
\change{2021-03-15}{Kevin Hammond}{FM (IOHK)}{Added Section on Design Notes to record discussion.}
\change{2021-04-01}{Kevin Hammond}{FM (IOHK)}{Added Appendix on Native Token Governance}
\change{2021-04-01}{Kevin Hammond}{FM (IOHK)}{Added Appendix on Expert Ballots}
\change{2021-04-09}{Kevin Hammond}{FM (IOHK)}{Added Section on Differentiated Submission Groups}
\change{2021-04-09}{Kevin Hammond}{FM (IOHK)}{Added Appendix on Efficiency Concerns}
\end{changelog}
