\pagebreak
\section{Expert Ballots}
\label{sect:expert-ballots}

This appendix considers how ``expert ballots'' could be integrated into the on-chain enactment process.
An expert ballot is a manifesto by a group of experts that presents an easily understood voting intention
(or collected group of intentions).  Under the assumption that it is not necessary to consider each issue
independently (which might place significant load on ada holders in assessing the quality of each ballot on
a regular basis), the PUP approach fits such an approach well.

\subsection{Proposal Submission}

No changes are needed to the proposal submission process.  Expert groups will monitor the progress of proposals
and update their expert ballots in advance of voting (perhaps even in advance of submission).

\subsection{Vote Delegation}

The delegate mechanism described here can be used to directly support expert
ballots.  Expert groups may register as delegates, and ada holders may choose to
delegate their vote to those groups either on a per-issue basis or over a period
of time.  As usual, Ada holders may change their voting intention at any point in time.
This flexible approach will suit enterprise ada holders (where it may take substantial
effort to change a set of vote delegations) as well as individuals (who may

\subsection{Endorsement and Enactment}

No changes are needed to the endorsement and enactment phases.  Protocol version updates will need
to be endorsed by block producers, and enactment can be carried out automatically.


\subsection{Voting Centre}

The voting centre would need to be changed so that expert ballots could be presented to ada holders,
and so that vote delegations could then be made to these expert groups.  Users might also want
to be alerted to new ballots (perhaps filtered in some way).

For example, a user might be presented with an expert ballot, given the opportunity to read the ballot
and conduct any background research, and then reject the ballot, select it as a voting choice, or
save it for later consideration.
