\hypersetup{
  pdftitle={Design of the Cardano Ledger with Automated Parameter Updates},
  breaklinks=true,
  bookmarks=true,
  colorlinks=false,
  linkcolor={blue},
  citecolor={blue},
  urlcolor={blue},
  linkbordercolor={white},
  citebordercolor={white},
  urlbordercolor={white}
}


\floatstyle{boxed}
\restylefloat{figure}
\cleardoublepage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}

\title{Design of the Cardano Ledger with Automated Parameter Updates and Central Fund Transfers}

\author{
   Kevin Hammond \\ {\small \texttt{kevin.hammond@iohk.io}} \\
   Philipp Kant \\ {\small \texttt{philipp.kant@iohk.io}} \\
   Andre Knispel \\ {\small \texttt{andre.knispel@iohk.io}} \\
   }

\date{}

\maketitle

\begin{abstract}
  This document outlines the design of the mechanisms that are needed to support
  automated parameter updates and central funds transfers.  This enables a key part of the Cardano governance structure by
  eliminating the use of genesis keys or delegates to control the operation of the blockchain.  The overall governance process includes both
  off-chain and on-chain components.  This design document is concerned primarily with the on-chain component.
\end{abstract}

\section*{List of Contributors}
\label{acknowledgements}

\begin{changelog}
\change{2021-02-16}{Kevin Hammond}{FM (IOHK)}{Initial version. }
\change{2021-02-17}{Kevin Hammond}{FM (IOHK)}{Description of goals and submission process.}
\change{2021-02-19}{Kevin Hammond}{FM (IOHK)}{Groups involved.  Vote delegation.}
\change{2021-02-19}{Kevin Hammond}{FM (IOHK)}{Added various outline sections.  Voting, sidechains, short Priviledge comparison.}
\end{changelog}
