\section{UTxO}
\label{sec:utxo}


\subsection{UTxO Transitions}
\label{sec:utxo-trans}

We have added several functions having to to with transaction and UTxO inputs and
outputs, which are used in defining the UTxO transition system. These are
given in Figure \ref{fig:functions:insouts}. These include

\begin{itemize}
  \item the function $\fun{txinputs_{vf}}$ returns only those transaction inputs
  that were selected to pay transaction fees (we call these "fee-marked" inputs)
  \item the predicate function $\fun{feesOK}$ checks whether the transaction is
  paying the necessary fees, and correctly. That is, it checks that
  \begin{itemize}
    \item[(i)] the fee-marked inputs are strictly of the type that are key or
    multisignature script locked, not Plutus script
    \item [(ii)] all the fee-marked inputs contain strictly Ada and no other tokens
    \item [(iii)] the fee-marked inputs are enough to cover the fee amount stated
    in the transaction
    \item [(iv)] minimum fee the transaction is obligated to pay (this includes
    the script-running fee) is less than the fee amount the transaction states
    it is paying
  \end{itemize}
  \item the function $\fun{getOut}$ selects from a transaction output the data that
  will be stored in the UTxO, i.e. $\UTxOOut$ without the slot number
  \item the function $\fun{outs}$ now adds the slot number of the block in which transaction is
  included to the outputs
  \item the function $\fun{txins}$ that returns UTxO output reference part of both script
  and non-script transaction inputs.
\end{itemize}

Note that when submitting a transaction, the wallet is responsible for
determining the total price of the
validation of all the Plutus scripts in a transaction
by running the script itself to see how much resources it takes and doing the
fee calculation using the cost model in protocol parameters. It is then
also responsible for adding enough inputs to the transaction to cover the
fees required.

\begin{figure}[htb]
  \begin{align*}
    & \fun{txinputs_{vf}} \in \TxBody \to \powerset{\TxId \times \Ix} \\
    & \text{tx VK inputs used for fees} \\
    & \fun{txinputs_{vf}} ~txb~= \\
    &~~\{ (txid,ix)~\vert~((txid,ix),\var{isfee}) \in
    \fun{txinputs} ~txb,~
     \var{isfee}\in\Yes\}
    \nextdef
    & \fun{feesOK} \in \N \to \PParams \to \GoguenTx \to \UTxO \to \Bool  \\
    & \text{check if fee-marked inputs are Ada-only and enough to cover fees} \\
    & \fun{feesOK} ~n~\var{pp}~tx~utxo~= \\
    &~~\fun{range}~(\fun{txinputs_{vf}}~{txb} \restrictdom \var{utxo}) \subseteq \TxOutND ~ \\
    &~~\wedge~ \fun{ubalance}~(\fun{txinputs_{vf}}~{txb}) \in \Coin \\
    &~~      \wedge~ \fun{ubalance}~(\fun{txinputs_{vf}}~{txb} \restrictdom \var{utxo}) \geq \txfee{txb} ~ \\
    &~~      \wedge~ \minfee~n~{pp}~{tx} \leq \txfee{txb} \\
    &~~      \where \\
    & ~~~~~~~ \var{txb}~=~\txbody{tx}
    \nextdef
    & \fun{getOut} \in \TxOut \to (\Addr \times \Value) \uniondistinct (\Addr \times \Value \times \DataHash)  \\
    & \text{tx outputs transformed to UTxO outputs} \\
    & \fun{getOut} ~{txout}~= \begin{cases}
         \var{txout}  & \text{if~} \var{txout} \in \TxOutND \\
              (\fun{getAddr}~\var{txout}, \fun{getValue}~\var{txout},
              \fun{getDataHash}~\var{txout}) & \text{otherwise}
            \end{cases}
    \nextdef
    & \fun{outs} \in \Slot \to \TxBody \to \UTxO \\
    & \text{tx outputs as UTxO} \\
    & \fun{outs} ~ \var{slot}~\var{txb} =
        \left\{
          (\fun{txid} ~ \var{txb}, \var{ix}) \mapsto (\fun{getOut}~\var{txout},\var{slot}) ~
          \middle|~
          \var{ix} \mapsto \var{txout} \in \txouts{txb}
        \right\} \\
    \nextdef
    & \fun{txins} \in \TxBody \to \powerset{\TxId \times \Ix} \\
    & \text{transaction inputs} \\
    & \fun{txins} ~\var{txb} = \{(txid,ix) \mid ((txid,ix),\wcard)\in\fun{txinputs} ~txb\} \\
  \end{align*}
  \caption{Functions on Tx Inputs and Outputs}
  \label{fig:functions:insouts}
\end{figure}


Figure~\ref{fig:functions:utxo} defines functions needed for the UTxO transition system.
The changes due to Plutus integration are as follows:

\begin{itemize}

  \item $\fun{getCoin}$ adds up all the Ada in a given output and returns it as a
  $\Coin$ value

  \item $\fun{utxoAda}$ returns the set of all the outputs in a UTxO with only Ada
  tokens (the other tokens are discarded). This is used in the stake distribution
  calculation at the epoch boundary

  \item $\fun{txscrfee}$ calculates the fee a transaction must pay for script
  execution based on current resource use pricing. Note that this value,
  like the non-script portion of the transaction fee, is calculated in $\Coin$,
  as fees can only be paid in Ada.

  \item For Goguen transactions, we have also changed the minimim fee calculation, $\fun{minfee}$,
  to include the script fees the transaction is obligated to pay to run its
  scripts.

  \item
    The $\fun{ubalance}$ function calculates the (aggregated by currency ID and
    Token) sum total of all the value in a given UTxO.

  \item The $\fun{consumed}$ calculation for the preservation of value remains the
  same. It is still the sum of the reward address value consumed, the value
   UTxO entries consumed, the value forged by a transaction,
   and the value consumed from the deposit pot due
   to the transaction collecting deposit refunds.

  \item The $\fun{produced}$ calculation contains the fee a transaction pays
  to the fee pot (this consists of the sum of two values, a transaction fee and a
  fee for processing all the scripts inside it), the outputs it adds to the UTxO,
  and the deposits it makes
  to the deposit pot (these are the same values as the ledger spec).
  The script execution costs are included in the transaction fee.
  This calculation now takes the current slot number as an argument.
\end{itemize}

\textbf{Value Operations and Partial Order.}
Note that
the $\fun{consumed}$ and $\fun{produced}$ values are compared as $\Value$.
The type $\Value$ has a partial order defined on it in (see~\cite{plutus_eutxo}).
The idea of this partial order, as well as operations like addition and subtraction
of terms of the $\Value$ type, is that a term $v \in \Value$ can be treated
as a total map that maps to the trivial element $\epsilon \in \Token \to \Quantity$
almost everywhere except for a several currency IDs that appear in $\dom~v$.

This way, when adding two terms of type value, $\{ cid1 \mapsto tkns1\}$ and
$\{ cid2 \mapsto tkns2\}$, the result is
\[ \{ cid1 \mapsto (tkns1 + \epsilon), cid2 \mapsto (\epsilon + tkns2)\} \]

which simplifies to

\[ \{ cid1 \mapsto tkns1, cid2 \mapsto tkns2\} \]

This definition of addition is also defined in Section~\ref{sec:notation-shelley}.
Similarly, if we compare these two tokens, we compare them as follows
(element-wise, with $\epsilon$ as the value associated to any currency ID
not present in on of the values),

\[ (\{ cid1 \mapsto tkns1\}~\leq~\{ cid2 \mapsto tkns2\}) \Leftrightarrow \\
   (tkns1 \leq \epsilon~\wedge~ \epsilon \leq tkns2) \]

\begin{figure}[htb]
  \emph{Helper Functions}
  \begin{align*}
    & \fun{getCoin} \in \UTxOOut \to \Coin \\
    & \fun{getCoin}~{\var{out}} ~=~\sum_{\mathsf{adaID} \mapsto tkns \in \fun{getValue}~out}
       (\sum_{q \in \range~{tkns}} \fun{co}~q) \\
    & \text{sum total of amount of Ada in an output}
    \nextdef
    & \fun{utxoAda} \in \UTxO \to \powerset{(\Addr \times \Coin)} \\
    & \fun{utxoAda}~{\var{utxo}} ~=~\{~(\fun{getAddr}~\var{out},~\fun{getCoin}~{out})
    ~\vert~ \var{out} \in \range~\var{utxo} ~\} \\
    & \text{returns the outputs in the UTxO with only the Ada coins} \\
  \end{align*}
  %
  \emph{Main Calculations}
  \begin{align*}
    & \fun{txscrfee} \in \N \to \Prices \to \ExUnits \to \Coin \\
    & \fun{txscrfee}~n~ (\var{pr_{init}, pr_{mem}, pr_{steps}})~ (\var{mem, steps})
    = \var{pr_{init}}*n + \var{pr_{mem}}*\var{mem} + \var{pr_{steps}}*\var{steps} \\
    & \text{calculates the script fee a transaction must pay} \\
    \nextdef
    &\fun{minfee} \in \N \to \PParams \to \GoguenTx \to \Coin \\
    & \text{minimum fee}\\
    &\fun{minfee}  ~n~\var{pp}~\var{tx} = \\
    &~~(\fun{a}~\var{pp}) \cdot \fun{txSize}~\var{tx} + (\fun{b}~\var{pp}) +
    \fun{txscrfee}~n~(\fun{prices}~{pp})~(\fun{txexunits}~(\fun{txbody}~{tx}))
    \nextdef
    & \fun{ubalance} \in \UTxO \to \Value \\
    & \fun{ubalance} ~ utxo = \sum_{\wcard\mapsto\var{(q,s)}\in~\var{utxo}}
    \fun{getValue}~\var{q} \\
    & \text{UTxO balance} \\
    \nextdef
    & \fun{consumed} \in \PParams \to \UTxO \to \StakeCreds \to \Wdrl \to \TxBody \to \Value \\
    & \consumed{pp}{utxo}{stkCreds}{rewards}~{txb} = \\
    & ~~\ubalance{(\txins{txb} \restrictdom \var{utxo})} + \\
    &~~  \fun{coinToValue}(\fun{wbalance}~(\fun{txwdrls}~{txb})~\\
        &~~+~ \keyRefunds{pp}{stkCreds}{txb}) +
        ~\fun{forge}~\var{txb} \\
    & \text{value consumed} \\
    \nextdef
    & \fun{produced} \in \Slot \to \PParams \to \StakePools \to \TxBody \to \Value \\
    & \fun{produced}~\var{slot}~\var{pp}~\var{stpools}~\var{txb} = \\
    &~~\ubalance{(\outs{slot}~{txb})}  + \fun{coinToValue}(\txfee{txb} \\
    &~~+ \deposits{pp}{stpools}~{(\txcerts{txb})})\\
    & \text{value produced} \\
  \end{align*}
  \caption{Functions used in UTxO rules}
  \label{fig:functions:utxo}
\end{figure}

\clearpage

\textbf{Putting Together Plutus Scripts and Their Inputs.}
In Figure \ref{fig:functions:script1} we give the helper functions needed to
retrieve all the data relevant to validation of Plutus scripts.
This includes,

\begin{itemize}
  \item $\fun{indexof}$ finds the index of a given certificate, value, input, or
  withdrawal in the list, finite map, or set of things of the corresponding type.
  This function assumes there is some ordering on each of these structures.
  This function is abstract because it assumes there is some ordering rather
  than giving it explicitly. The specific ordering of a set or a finite map
  could be implementation-dependent. A list ordering should be unambiguous.
  \item $\fun{indexedScripts}$ creates a finite map wherein all the scripts
  of a transaction (that it carries as a set) are indexed by their hashes,
  calculated by this function
  \item $\fun{indexedDats}$ is a similar function that indexes all the datum objects
  carried by the transaction with their hashes
  \item $\fun{getmdl}$ selects, given a script and protocol parameters,
  the cost model that is associated with the language of that
  script in the protocol parameters
  \item $\fun{findRdmr}$ gets the redeemer carried by a Goguen transaction
   which corresponds to the given current item. Recall that items that may require
   Plutus scripts to be run include certificates, withdrawals, forges,
   and outputs. To find the redeemer corresponding to that item, we search
   the indexed redeemer structure for the redeemer with the right key.
   The key we look for is the
   pair of the type of item it is for (indicated by the tag), and the index of
   said item in the list/set/map of these kinds of items in the transaction.
\end{itemize}


\begin{figure}[htb]
  %
  \emph{Abstract functions}
  \begin{align*}
    &\fun{indexof} \in \DCert \to \seqof{\DCert} \to  \to \Ix\\
    &\fun{indexof} \in \Wdrl \to \Wdrl \to \Ix\\
    &\fun{indexof} \in \TxIn \to \powerset{\TxIn} \to \Ix\\
    &\fun{indexof} \in \Value \to \Value \to \Ix\\
    & \text{get the index of an item in the an ordered representation} \\
    \nextdef
  \end{align*}
  %
  \emph{Helper functions}
  \begin{align*}
    &\fun{indexedScripts} \in \GoguenTx \to (\ScriptHash \mapsto \Script) \\
    & \text{make a finite map of hash-indexed scripts} \\
    &\fun{indexedScripts}~{tx} ~=~ \{ h \mapsto s ~\vert~ \fun{hashScript}~{s}~=~h,
     s\in~\fun{txscripts}~(\fun{txwits}~{tx})\}
    \nextdef
    &\fun{indexedDats} \in \GoguenTx \to (\DataHash \mapsto \Data)\\
    & \text{make a finite map of hash-indexed datum objects} \\
    &\fun{indexedDats}~{tx} ~=~ \{ h \mapsto d ~\vert~ \fun{hashData}~{d}~=~h,
     d\in~\fun{txdats}~(\fun{txwits}~{tx})\}
    \nextdef
    &\fun{getmdl} \in \Script \to \PParams \to \CostMod \\
    & \text{make a finite map of hash-indexed scripts} \\
    &\fun{getmdl}~{s}~{pp} = \fun{head}~ (\fun{toList}~ (\range ((\fun{language}~{s})\restrictdom(\fun{costmdls}~{pp}))))
    &\nextdef
    &\fun{findRdmr} \in \GoguenTx \to \CurItem \to \powerset{\Data}\\
    & \text{get empty set or redeemer corresponding to index} \\
    & \fun{findRdmr}~{tx}~\var{it} ~=~ \{~ r ~\vert~ \\
    &~~(\mathsf{certTag}, \fun{indexof}~\var{it}~(\fun{txcerts}~{txb})) \mapsto ~r \in \fun{txrdmrs}~{txw} \\
    &~~\vee~ (\mathsf{wdrlTag}, \fun{indexof}~\var{it}~(\fun{txwdrls}~{txb}))
      \mapsto ~r \in \fun{txrdmrs}~{txw} \\
    &~~\vee~(\mathsf{forgeTag}, \fun{indexof}~\var{it}~(\fun{forge}~{txb}))  \mapsto ~r
      \in \fun{txrdmrs}~{txw} \\
    &~~\vee~(\mathsf{inputTag}, \fun{indexof}~\var{it}~(\fun{txinputs}~{txb})) \mapsto ~r
      \in \fun{txrdmrs}~{txw} \} \\
      & ~~\where \\
      & ~~~~~~~ \var{txb}~=~\txbody{tx} \\
      & ~~~~~~~ \var{txw}~=~\fun{txwits}~{tx}
  \end{align*}
  \caption{Combining Script Validators and their Inputs}
  \label{fig:functions:script1}
\end{figure}

\textbf{Matching Scripts and Inputs.}
In Figures \ref{fig:functions:script2} and \ref{fig:functions:script3},
we give the four functions that gather all data inside a transaction
and in the UTxO that is needed for script validation.

\begin{itemize}
\item $\fun{allCertScrts}$ returns the set of all the validators for the
key deregistration certificates, together
with the data needed for validation
\item $\fun{allWDRLSScrts}$ returns the set of all the validators locking
the script-address reward addresses together
with the data needed for validation
\item $\fun{forgedScrts}$ returns the set of all the validators for
forging new tokens together
with the data needed for validation
\item $\fun{allInsScrts}$ returns the set of all the validators locking
the script-address UTxO's together
with the data needed for validation
\end{itemize}


\begin{note}
\textbf{What scripts get redeemers?}
Here we assume it is ok for every kind of Plutus script to
have a redeemer. In fact, the transaction is obligated to provide a redeemer for
every Plutus script. There is the possibility, in the future, to optimize supplying redeemers
by allowing transactions to omit unit-value redeemers, filling the $\Data$-type
unit in by default. Whether this will be done is contingent on real-world
observations of the use of redeemers.
\end{note}

Note that there are no "checks" done inside these functions. If there are
missing validators or inputs, or incorrect hashes, wrong type of script,
this is caught during the application of the UTXOW
rule (before these functions are ever applied).
There are several pieces of data from different sources involved in building these
sets:

\begin{itemize}
\item the hash of the validator script, which is either the address (withdrawal
address or an output address in the UTxO),
the certificate credential, or the currency ID of forged tokens

\item the corresponding
full validator, which is looked up (by hash value) in the finite map
constructed by $\fun{indexedScripts}$

\item the datum objects, which are also looked up by hash in the map
constructed by $\fun{indexedDats}$.
The hashes used to look up the datum objects are found in the outputs of the UTxO,
indexed by the $(txid,ix)$ in the transaction output spending the UTxO entry

\item the redeemers, which are in the indexed redeemer structure carried by the transaction.
These are looked up by current item using the $\fun{findRdmr}$ function.

\item the validation data, built using the UTxO, the transaction itself,
and the current item being validated
\end{itemize}

Recall that $\fun{valContext}$ constructs the validation context (kept abstract
in this spec).

The function $\fun{mkPLCLst}$ returns a list made of the union of the sets of
pairs (of a script and the input list) which were constructed by the
functions specific to the script uses ($\fun{allCertScrts}$, $\fun{allWDRLSScrts}$,
$\fun{forgedScrts}$, $\fun{allInsScrts}$). Note that like the functions building
the script-inputs sets, $\fun{mkPLCLst}$ does not include the execution units or cost model
 parameters in its return type. This data is supplied to the Plutus interpreter
 directly during the application of the validation rule SVAL.

\begin{figure}[htb]
  \begin{align*}
    & \fun{allCertScrts} \in \UTxO \to \GoguenTx \to \powerset{(\ScriptPlutus \times \seqof{\Data})} \\
    & \text{check that all certificate witnessing scripts in a tx validate} \\
    & \fun{allCertScrts}~{utxo}~{tx}~=~ \\
    & ~~\{ (\var{script_v}, r;
    \fun{valContext}~\var{utxo}~\var{tx}~\var{cert}; \epsilon) ~\vert \\
    & ~~r \in \fun{findRdmr}~{tx}~\var{cert}, \\
    & ~~\var{cert} \in (\DCertDeRegKey\cap\fun{txcerts}~\txbody{tx}), \\
    &~~\fun{regCred}~\var{cert}\mapsto \var{script_v}\in \fun{indexedScripts}~{tx}, \\
    & ~~\var{script_v} \in \ScriptPlutus
     \}
    %
    \nextdef
    & \fun{allWDRLSScrts} \in \UTxO \to \GoguenTx \to \powerset{(\ScriptPlutus\times\seqof{\Data})} \\
    & \text{check that all reward withdrawal locking scripts in a tx validate} \\
    & \fun{allWDRLSScrts}~{utxo}~{tx}~=~ \\
    & ~~\{ (\var{script_v}, r; \fun{valContext}~\var{utxo}~\var{tx}~
      (a\mapsto c); \epsilon) ~\vert \\
    &~~ r\in \fun{findRdmr}~{tx}~\var{(a\mapsto c)}, \\
    & ~~a \mapsto c \in\fun{txwdrls}~(\txbody{tx}), \\
    & ~~\fun{validatorHash}~\var{a}\mapsto \var{script_v}\in \fun{indexedScripts}~{tx}, \\
    & ~~ \var{script_v} \in \ScriptPlutus \}
    \nextdef
    %
    & \fun{forgedScrts} \in \UTxO \to \GoguenTx \to \powerset{(\ScriptPlutus\times\seqof{\Data})} \\
    & \text{check that all forging scripts in a tx validate} \\
    & \fun{forgedScrts}~{utxo}~{tx}~=~\\
    & ~~\{ (\var{script_v}, r;
    \fun{valContext}~\var{utxo}~\var{tx}~\var{cid}; \epsilon) ~\vert \\
    & ~~r \in \fun{findRdmr}~{tx}~(\var{cid}\mapsto ~ \var{tkns}), \\
    & ~~\var{cid}\mapsto ~ \var{tkns} \in \fun{forge}~(\txbody{tx}), \\
    &~~\var{cid}\mapsto \var{script_v}\in \fun{indexedScripts}~{tx}
     \}
    %
  \end{align*}
  \caption{Scripts and their Arguments}
  \label{fig:functions:script2}
\end{figure}

\begin{figure}[htb]
  \begin{align*}
    & \fun{allInsScrts} \in \UTxO \to \GoguenTx \to \powerset{(\ScriptPlutus\times\seqof{\Data})} \\
    & \text{check that all UTxO entry locking scripts in a tx validate} \\
    & \fun{allInsScrts}~{utxo}~{tx}~=~ \{ (\var{script_v}; (\var{d};\var{r}; \\
    & ~~ \fun{valContext}~\var{utxo}~\var{tx}~
      (txid,ix,\var{hash_r}))) ~\vert \\
    & ~~(txid,ix) \in \fun{txinputs}~(\txbody{tx}), \\
    & ~~\var{r} \in \fun{findRdmr}~{tx}~\var{(txid,ix)}, \\
    & ~~\var{(txid,ix)} \mapsto ((a,v),h_d) \in \var{utxo}, \\
    & ~~\var{h_d}\mapsto \var{d} \in \fun{indexedDats}~{tx}, \\
    & ~~(\fun{validatorHash}~{a})\mapsto \var{script_v}\in \fun{indexedScripts}~{tx} \}
    \nextdef
    & \fun{mkPLCLst} \in \GoguenTx \to \UTxO \to \seqof{(\ScriptPlutus \times \seqof{\Data})} \\
    & \text{a list of all Plutus validators and corresponding input data} \\
    & \fun{mkPLCLst} ~\var{tx}~ \var{utxo} ~=~
    \fun{toList}~(\fun{allCertScrts}~{utxo}~{tx} \cup \fun{allWDRLSScrts}~{utxo}~{tx} \\
    & ~~ \cup \fun{allInsScrts}~{utxo}~{tx} ~~ \cup \fun{forgedScrts}~{utxo}~{tx}) \\
  \end{align*}
  \caption{Scripts and their Arguments}
  \label{fig:functions:script3}
\end{figure}

\textbf{Two Phase Script Validation.}
Two phase Plutus script validation is necessary to ensure users pay for the
computational resources script validation uses.
Native script execution costs are expected to be much smaller than Plutus
scripts, and can be assesed and limited by the ledger rules directly.
Hence these scripts do not require two-phase validation. They are already
in use in the Shelley spec with a signle validation phase.

The first phase two-phase validation approach
performs every aspect of transaction validation except running the scripts.
The second phase is running the scripts. We use four transition systems
for this validation approach, each with different responsibilities. We
give the details of each below, but to summarize, when a transction is processed,
it is done by rules in the transition systems in the following order
(each transition calls on the one below it in its rules):

\begin{itemize}
  \item[(UTXOW)] : Verifies all the necessary witnessing info is present, including
  VK witnesses, scripts, and all the script input data. It also performs
  key witness checks and runs multisig scripts. It then applies state changes
  computed by the UTXO transition
  \item[(UTXO)] : Verifies a transaction satisfies all the accounting requirements
  (including the general accounting property, correct fee payment, etc.),
  applies state changes computed by the UTXOS transition
  \item[(UTXOS)] : Performs the appropriate UTxO state changes, deciding based on the
  value of the $\IsValidating$ tag, which it checks using the SVAL transition
  \item[(SVAL)] : Runs the scripts, verifying that the $\IsValidating$ tag
  is applied correctly
\end{itemize}


Recall that, unlike native
multisignature scripts, Plutus scripts are opaque to the ledger. Recall also
that a transaction states a $\ExUnits$ "budget" to cover running all Plutus
scripts it is carrying. There is no way to check that this budget is enough,
except running the scripts. To avoid over-spending, we run them sequentially,
stopping whenever one does not validate, and charging the transaction the
fees. From the point of view of the ledger, there is no difference
between a script runnig out of $\ExUnits$ during validation, or not validating.
If a transaction contains an invalid script, the only change to the ledger
as a result of applying this transaction is the fees. Other parts of
the transaction cannot be processed correctly in this case.

Two phase validation requires a new transition system
(see Figure \ref{fig:ts-types:utxos}) to sequentially run
scripts and keep track of the execution units being spent as part of its state
($\var{remExU}$). The signal here is a sequence of pairs of a validator
script and the corresponding input data.

Note that there is one state variable in the SVAL transition system. The reason
for this is that in the second, script-running validation phase, we separate
the UTxO state update from sequentially running scripts. This transition
system is strictly for running the scripts, and a transition of this type
will be used by another rule to perform the correct UTxO update.

Running scripts sequentially
to verify that they all validate in the alloted $\ExUnits$ budget only requires
the amount of remaining $\ExUnits$ to be included in the state, and nothing else.
In the environment, we need the protocol parameters and the
transaction being validated. All other data needed
to run the scripts comes from the signal.

\begin{figure}[htb]
  \emph{Validation environment}
  \begin{equation*}
    \ValEnv =
    \left(
      \begin{array}{r@{~\in~}lr}
        \var{pp} & \PParams & \text{protocol parameters}\\
        \var{tx} & \GoguenTx & \text{transaction being processed} \\
      \end{array}
    \right)
  \end{equation*}
  %
  \emph{Validation state}
  \begin{equation*}
    \ValState =
    \left(
      \begin{array}{r@{~\in~}lr}
        \var{remExU} & \ExUnits & \text{exunits remaining to spend on validation} \\
      \end{array}
    \right)
  \end{equation*}
  %
  \emph{Script transitions}
  \begin{equation*}
    \_ \vdash
    \var{\_} \trans{sval}{\_} \var{\_}
    \subseteq \powerset (\ValEnv \times \ValState \times \seqof{(\ScriptPlutus\times\seqof{\Data})} \times \ValState)
  \end{equation*}
  %
  \caption{UTxO script validation types}
  \label{fig:ts-types:utxos}
\end{figure}

The rules for the second-phase script validation SVAL are given in
Figure~\ref{fig:rules:utxo-scrval}. Again, there is no UTxO state update
done in this rule. Its function is essentially verifying that the
validation tag ($\fun{txvaltag}$) is applied correctly by the creater of
the block by running all the scripts.

Note that following the Shelley ledger spec approach, every function
we define and use in the preconditions or calculations in the rules is
necessarily total.
This way, all errors (validation failures) we encounter always come from
rule applications, i.e. a precondition of a rule is not met.
We mention this here because the SVAL rule looks as if it could be
simply a function. However, we want the incorrect application of the
validation tag to be an error, so it must be an error that comes form
an unmet precondition of a rule.

There are three transition rules.
The first rule, $\mathsf{Scripts\mbox{-}Val}$, applies when

\begin{itemize}
\item there
are no scripts left to validate in the signal list (i.e. this is the base case of
induction when all the scripts have validated)
\item the validation tag is applied correctly (it is $\Yes$)
\end{itemize}

The $\mathsf{Scripts\mbox{-}Stop}$ rule applies when

\begin{itemize}
  \item The currenct script-input pair being validated does not validate
  (because there are not enough $\ExUnits$ left or any other reasons)
  \item The validation tag is correct ($\Nope$ in this case)
\end{itemize}

These first two rules require no state change.
The $\mathsf{Scripts\mbox{-}Ind}$ rule applies when

\begin{itemize}
  \item the current script being validated has validated
  \item there is a non-negative fee which remains to pay for validating
  the rest of the scripts in the list
  \item transition rules apply for rest of the list (without the currenct script)
\end{itemize}

The only state change in this rule is of the variable $\var{remExU}$.
It is decreased by subtracting the cost of the execution of the
current script from its current value.
This is the variable we use to keep track of the remaining funds for
script execution. If the transaction is overpaying ($\fun{txscrfee}~{tx}$
is too big), the whole fee is still taken.

It is always in the interest of the slot leader to have the new block validate,
containing only valid transactions. This motivates the
slot leader to:

\begin{itemize}
  \item correctly apply of the $\IsValidating$ tag,
  \item include transactions that validate in every way
  \textit{except possibly 2nd step script validation failure}
  \item exclude any transactions that are invalid
in some way \textit{other than 2nd step script validation failure}
\end{itemize}

We want to
throw away all the blocks which have transactions with these tags
applied incorrectly.
One of the reasons for having the correct validation tag added by the slot leader
to a transaction is that re-applying blocks would not require repeat
execution of scripts in the transactions inside a block. In fact, when replaying
blocks, all the witnessing info can be thrown away.
We also rely on correct use of tags in other rules (at this time, only in
the rules in Figure \ref{fig:rules:ledger}).

\begin{note}
\textbf{Non-integral calculations inside the Plutus interpreter:} if there
will be some in the future (from the Actus contracts implemented using
the Marlowe interpreter, for e.g.) they should
be done the same way they are done in the Shelley ledger. This is a matter of
deterministic script validation outcomes. Inconsistent rounding could
result in different validation outcomes running the same script on the same
arguments. For how this is done in the ledger calculations, see
~\cite{non_int}.
\end{note}


\begin{figure}[htb]
  \begin{equation}
    \inference[Scripts-Val]
    {
    \fun{txvaltag}~\var{tx} \in \Yes  &
    \var{remExU}~\geq~0
    }
    {
    \begin{array}{l}
      \var{pp}\\
      \var{tx}\\
    \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      \right)
      \trans{sval}{\epsilon}
      \left(
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      \right) \\
    }
  \end{equation}
  \begin{equation}
    \inference[Scripts-Stop]
    { \\~\\
    (\var{isVal},\var{remExU'})~:=~ \llbracket \fun{fst}~{s} \rrbracket_
    {\fun{getmdl}~(\fun{fst}~{s})~\var{pp},\var{remExU}} (\fun{snd}~{s})
    \\
    ~
    \\
    \fun{txvaltag}~\var{tx} \in \Nope &
    (\var{remExU'}~<~0 ~ \lor ~ \var{isVal}\in \Nope)
    }
    {
    \begin{array}{l}
      \var{pp}\\
      \var{tx}\\
    \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      \right)
      \trans{sval}{\Gamma;s}
      \left(
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      \right)
    }
  \end{equation}
  \begin{equation}
    \inference[Scripts-Ind]
    { \\~\\
    (\var{isVal},\var{remExU''})~:=~ \llbracket \fun{fst}~{s} \rrbracket
    _{\fun{getmdl}~(\fun{fst}~{s})~\var{pp},\var{remExU'}} (\fun{snd}~{s})
    \\
    ~
    \\
    \var{remExU''}~\geq~0
    \\
    ~
    \\
    {
    \begin{array}{l}
      \var{pp}\\
      \var{tx}\\
    \end{array}
    }
      \vdash
      \left(
      {
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      }
      \right)
      \trans{sval}{\Gamma}
      \left(
      {
      \begin{array}{r}
        \var{remExU'}\\
      \end{array}
      }
      \right)
    }
    {
    \begin{array}{l}
      \var{pp}\\
      \var{tx}\\
    \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{remExU}\\
      \end{array}
      \right)
      \trans{sval}{\Gamma;s}
      \left(
      \begin{array}{r}
        \varUpdate{remExU''}\\
      \end{array}
      \right)
    }
  \end{equation}
  \caption{Script validation rules}
  \label{fig:rules:utxo-scrval}
\end{figure}


\textbf{Updating the UTxO State.}

We have defined a separate transition system, UTXOS, to represent the two distinct
UTxO state changes, one resulting from all scripts in a transaction validating,
the other - from at least one failing to validate. Its transition types
are all the same as for the for the UTXO transition, see Figure
\ref{fig:ts-types:utxo-scripts}.

\begin{figure}[htb]
  \emph{State transitions}
  \begin{equation*}
    \_ \vdash
    \var{\_} \trans{utxos}{\_} \var{\_}
    \subseteq \powerset (\UTxOEnv \times \UTxOState \times \GoguenTx \times \UTxOState)
  \end{equation*}
  %
  \caption{UTxO state update types}
  \label{fig:ts-types:utxo-scripts}
\end{figure}

There are two rules corresponding to the two possible state changes of the
UTxO state in the UTXOS transition system, see Figure~\ref{fig:rules:utxo-state-upd}.

In both cases, the SVAL transition is called upon to verify that the $\IsValidating$
tag has been applied correctly. The function $\fun{mkPLCLst}$ is used to build
the signal list $\var{sLst}$ for the SVAL transition.

The first rule
applies when the validation tag is $\Yes$.
In this case, the states of the UTxO, fee
  and deposit pots, and updates are updated exactly as in the current Shelley
  ledger spec.

  The second rule
  applies when the validation tag is $\Nope$.
  In this case, the UTxO state changes as follows:

  \begin{itemize}
    \item All the
    UTxO entries corresponding to the transaction inputs selected for covering
    script fees are removed

    \item The sum total of the value of the marked UTxO entries
    is added to the fee pot
  \end{itemize}


\begin{figure}[htb]
  \begin{equation}
    \inference[Scripts-Yes]
    {
    \var{txb}\leteq\txbody{tx} &
    \fun{txvaltag}~\var{tx} \in \Yes
    \\
    ~
    \\
    \var{sLst} := \fun{mkPLCLst}~\var{tx}~\var{utxo}
    \\~\\
    {
      \left(
        \begin{array}{r}
          \var{pp} \\
          \var{tx} \\
        \end{array}
      \right)
    }
      \vdash
        \var{\fun{txexunits}~{tx}}
      \trans{sval}{sLst}\var{remExU}
      \\~\\
    {
      \left(
        \begin{array}{r}
          \var{slot} \\
          \var{pp} \\
          \var{genDelegs} \\
        \end{array}
      \right)
    }
    \vdash \var{ups} \trans{\hyperref[fig:rules:update]{up}}{\fun{txup}~\var{tx}} \var{ups'}
    \\~\\
    \var{refunded} \leteq \keyRefunds{pp}{stkCreds}~{txb}
    \\
    \var{decayed} \leteq \decayedTx{pp}{stkCreds}~{txb}
    \\
    \var{depositChange} \leteq
      (\deposits{pp}~{stpools}~{\txcerts{txb}}) - (\var{refunded} + \var{decayed})
    }
    {
    \begin{array}{l}
      \var{slot}\\
      \var{pp}\\
      \var{stkCreds}\\
      \var{stpools}\\
      \var{genDelegs}\\
    \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{utxo} \\
        \var{deposits} \\
        \var{fees} \\
        \var{ups} \\
      \end{array}
      \right)
      \trans{utxos}{tx}
      \left(
      \begin{array}{r}
        \varUpdate{\var{(\txins{txb} \subtractdom \var{utxo}) \cup \outs{slot}~{txb}}}  \\
        \varUpdate{\var{deposits} + \var{depositChange}} \\
        \varUpdate{\var{fees} + \txfee{txb} + \var{decayed}} \\
        \varUpdate{\var{ups'}} \\
      \end{array}
      \right) \\
    }
  \end{equation}
  \begin{equation}
    \inference[Scripts-No]
    {
    \var{txb}\leteq\txbody{tx} &
    \fun{txvaltag}~\var{tx} \in \Nope
    \\
    ~
    \\
    \var{sLst} := \fun{mkPLCLst}~\var{tx}~\var{utxo}
    \\~\\
    {
      \left(
        \begin{array}{r}
          \var{pp} \\
          \var{tx} \\
        \end{array}
      \right)
    }
      \vdash
        \var{\fun{txexunits}~{tx}}
      \trans{sval}{sLst}\var{remExU}
    }
    {
    \begin{array}{l}
      \var{slot}\\
      \var{pp}\\
      \var{stkCreds}\\
      \var{stpools}\\
      \var{genDelegs}\\
    \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{utxo} \\
        \var{deposits} \\
        \var{fees} \\
        \var{ups} \\
      \end{array}
      \right)
      \trans{utxos}{tx}
      \left(
      \begin{array}{r}
        \varUpdate{\var{\fun{txinputs_{vf}}~{txb} \subtractdom \var{utxo}}}  \\
        \var{deposits} \\
        \varUpdate{\var{fees} + \fun{ubalance}~(\fun{txinputs_{vf}}~{txb}\restrictdom \var{utxo})} \\
        \var{ups} \\
      \end{array}
      \right)
    }
  \end{equation}
  \caption{State update rules}
  \label{fig:rules:utxo-state-upd}
\end{figure}

In Figure \ref{fig:rules:utxo-shelley}, we present the $\type{UTxO-inductive}$
transition rule for the UTXO transition type. Note that the
signal for this transition is now specifically of type $\GoguenTx$, it does not
work with Shelley transactions (see explanation about transforming one type into
the other below). This rule
It has the following preconditions (the relevant ones remain
from the original Shelley spec):

\begin{itemize}
  \item The transaction is being processed within its validity interval

  \item The transaction has at least one input

  \item All inputs in a transaction correspond to UTxO entries

  \item The general accounting property holds

  \item The transaction is paying fees correctly

  \item The transaction is not forging any Ada

  \item All outputs of the transaction contain only non-negative quantities

  \item The transaction size does not exceed maximum

  \item The execution units budget a transaction gives does not exceed the max
  allowed units

  \item The UTXOS state transition is valid
\end{itemize}

The resulting state transition is defined entirely by the application of the
UTXOS rule.

\begin{figure}[htb]
  \begin{equation}\label{eq:utxo-inductive-shelley}
    \inference[UTxO-inductive]
    {
      \var{txb}\leteq\txbody{tx} &
      \var{txw}\leteq\fun{txwits}~{tx} \\
      \fun{txfst}~txb \leq \var{slot}
      & \fun{txttl}~txb \geq \var{slot}
      \\
      \txins{txb} \neq \emptyset
      & \txins{txb} \subseteq \dom \var{utxo}
      \\
      \consumed{pp}{utxo}{stkCreds}{rewards}~{txb} = \produced{slot}~{pp}~{stpools}~{txb}
      \\~\\
      \fun{feesOK}~(\vert~ \fun{txscripts}~{tx} \cap \ScriptPlutus ~\vert) ~pp~tx~utxo \\
      \\
      ~
      \\
      \mathsf{adaID}~\notin \dom~{\fun{forge}~tx} \\
      \forall txout \in \txouts{txb}, ~ \fun{getValue}~txout\}  ~\geq ~ \epsilon \\~
      \\
      \fun{txsize}~{tx}\leq\fun{maxTxSize}~\var{pp} \\
      \fun{txexunits}~{txb} \leq \fun{maxTxExUnits}~{pp}
      \\
      ~
      \\
      {
        \begin{array}{c}
          \var{slot}\\
          \var{pp}\\
          \var{stkCreds}\\
          \var{stpools}\\
          \var{genDelegs}\\
        \end{array}
      }
      \vdash
      {
        \left(
          \begin{array}{r}
            \var{utxo} \\
            \var{deposits} \\
            \var{fees} \\
            \var{ups}\\
          \end{array}
        \right)
      }
      \trans{utxos}{\var{tx}}
      {
        \left(
          \begin{array}{r}
            \var{utxo'} \\
            \var{deposits'} \\
            \var{fees'} \\
            \var{ups'}\\
          \end{array}
        \right)
      }
    }
    {
      \begin{array}{l}
        \var{slot}\\
        \var{pp}\\
        \var{stkCreds}\\
        \var{stpools}\\
        \var{genDelegs}\\
      \end{array}
      \vdash
      \left(
      \begin{array}{r}
        \var{utxo} \\
        \var{deposits} \\
        \var{fees} \\
        \var{ups}\\
      \end{array}
      \right)
      \trans{utxo}{tx}
      \left(
      \begin{array}{r}
        \varUpdate{\var{utxo'}}  \\
        \varUpdate{\var{deposits'}} \\
        \varUpdate{\var{fees'}} \\
        \varUpdate{\var{ups'}}\\
      \end{array}
      \right)
    }
  \end{equation}
  \caption{UTxO inference rules}
  \label{fig:rules:utxo-shelley}
\end{figure}

\clearpage

\textbf{Witnessing.}

Plutus script validation is not part of witnessing because of the introduction
of two-phase validation, as this type of validation may result in two
different ways of updating the UTxO (fee payment only, or a full update).
Native script validation
still is, and we need to pick only the native scripts to validate as
part of witnessing. We have changed the definition of the function
$\fun{scriptsNeeded}$, see Figure~\ref{fig:functions-witnesses}.

The result type of this function is now a finite map that associates items
that require script validation with the hash of the script that must be run.
This includes both MSig and Plutus scripts, and scripts used for every
validation purpose (forging, outputs, certificates, withdrawals), see Figure
\ref{fig:functions-witnesses}.


\begin{figure}[htb]
  \begin{align*}
      & \hspace{-1cm}\fun{scriptsNeeded} \in \UTxO \to \GoguenTx \to
        (\CurItem \mapsto \ScriptHash)\\
      & \hspace{-1cm}\text{items that need script validation and corresponding script hashes} \\
      &  \hspace{-1cm}\fun{scriptsNeeded}~\var{utxo}~\var{tx} = \\
      & ~~\{ i \mapsto \fun{validatorHash}~(\fun{getAddr}~{txout}) \mid i \mapsto \var{txout} \in \var{utxo},\\
      & ~~~~~i\in\fun{txinsScript}~{(\fun{txinputs}~\var{txb})}~{utxo}\} \\
      \cup & ~~\{ (a\mapsto c) \mapsto \fun{stakeCred_{r}}~\var{a} \mid a \mapsto c \in \fun{txwdrls}~\var{txb}),
         a\in \AddrRWDScr \} \\
        \cup & ~~\{ cert \mapsto a \mid a \in\ScriptHash \cap \fun{certWitsNeeded}~{txb}\} \\
        \cup & ~~\{ (cid \mapsto \var{tkns}) \mapsto \var{cid} \mid cid \mapsto \var{tkns}~\in~\fun{forge}~{txb} \} \\
      & \where \\
      & ~~~~~~~ \var{txb}~=~\txbody{tx} \\
      & ~~~~~~~ \var{txw}~=~\fun{txwits}~{tx}
  \end{align*}
  \caption{Functions used in witness rule}
  \label{fig:functions-witnesses}
\end{figure}

Recall here that in the Goguen era, we must be able to validate both Shelley type
and Goguen type transactions. To do this, we have two separate rules in for the
UTXOW transition type. The first rule applies when we have $tx \in \GoguenTx$, see
Figure~\ref{fig:rules:utxow-goguen}. In addition to the Shelley UTXOW preconditions
that still apply, we have made the following changes and additions to the
preconditions:

\begin{itemize}

    \item All the multisig scripts the transaction is carrying validate

    \item The transaction has exactly the scripts required for witnessing and no
    additional ones (this includes all languages of scripts, for all purposes)

    \item The transaction is carrying a redeemer for every item that needs
    validation by a Plutus script

    \item The only certificates that are allowed to have scripts as witnesses
    are delegation deregistration certificates

    \item The transaction has a datum for every Plutus script output it is spending

    \item The transaction has a datum for every Plutus script output that is
    marked with the $\Yes$ tag for $\HasDV$

    \item
    The hash of the subset of protocol parameters in the transaction body is equal to
    the hash of the same subset of protocol parameters currently on the ledger

    \item The hash of the indexed redeemer structure attached to the transaction is
    the same as the $\fun{rdmrsHash}~{tx}$ (the hash value contained in the signed body of
    the transaction)

\end{itemize}

If these conditions are all satisfied, the resulting UTxO state change is fully determined
by the UTXO transition (the application of which is also part of the conditions).

\begin{figure}
  \begin{equation}
    \label{eq:utxo-witness-inductive-goguen}
    \inference[UTxO-witG]
    {
      \var{txb}\leteq\txbody{tx} &
      \var{txw}\leteq\fun{txwits}~{tx} &
      \var{tx}~\in~\GoguenTx \\
      (utxo, \wcard, \wcard, \wcard) \leteq \var{utxoSt} \\
      \var{witsKeyHashes} \leteq \{\fun{hashKey}~\var{vk} \vert \var{vk} \in
      \dom (\txwitsVKey{txw}) \}\\~\\
      \forall \var{validator} \in \fun{txscripts}~{txw} \cap \ScriptMSig,\\
      \fun{runMSigScript}~\var{validator}~\var{tx}\\~\\
      \range~(\fun{scriptsNeeded}~\var{utxo}~\var{tx}) ~=~ \dom (\fun{indexedScripts}~{tx}) \\
      \forall c\mapsto h \in ~\fun{scriptsNeeded}~\var{utxo}~\var{tx}, ~h\mapsto s~\in~\fun{indexedScripts}~{tx},\\
       s \in \ScriptPlutus~\Leftrightarrow ~\fun{findRdmr}~{tx}~{c}\neq \emptyset
      \\~\\
      \forall \var{cert}~\in~\fun{txcerts}~{txb}, \fun{regCred}~{cert}\in \ScriptHash \Leftrightarrow
      \var{cert} \in~ \DCertDeRegKey \\~\\
      \forall~\var{txin}\in\fun{txinputs}~{txb},
      \var{txin} \mapsto \var{(\wcard,\wcard,h_d)} \in \var{utxo},
      \var{h_d} ~\in \fun{dom}(\fun{indexedDats}~{txw})
      \\
      ~
      \\
      \forall~ix \mapsto (a,v,d_h,\Yes) ~\in~\fun{txouts}~{txb}, \\
       \var{d_h}\in \fun{dom}~ (\fun{indexedDats}~{txw})
      \\
      ~
      \\
      \fun{ppHash}~{txb}~=~\fun{hashLanguagePP}~\var{pp}~(\fun{cmlangs}~(\fun{txscripts}~\var{txw})) \\~\\
      \fun{txrdmrs}~\var{txw} ~=~ \emptyset \Leftrightarrow \fun{rdmrsHash}~{txb}~=~\Nothing \\
      \fun{txrdmrs}~\var{txw} ~\neq~ \emptyset \Leftrightarrow
      \fun{hash}~(\fun{txrdmrs}~\var{txw})~ =~  \fun{rdmrsHash}~{txb} \\
      \\~\\
      \forall \var{vk} \mapsto \sigma \in \txwitsVKey{txw},
      \mathcal{V}_{\var{vk}}{\serialised{\txbody{tx}}}_{\sigma} \\
      \fun{witsVKeyNeeded}~{utxo}~{tx}~{genDelegs} \subseteq \var{witsKeyHashes}
      \\~\\
      genSig \leteq
      \left\{
        \fun{hashKey}~gkey \vert gkey \in\dom{genDelegs}
      \right\}
      \cap
      \var{witsKeyHashes}
      \\
      \left\{
        c\in\txcerts{txb}~\cap\DCertMir
      \right\} \neq\emptyset \implies \vert genSig\vert \geq \Quorum \wedge
      \fun{d}~\var{pp} > 0
      \\~\\
      \var{mdh}\leteq\fun{txMDhash}~\var{txb}
      &
      \var{md}\leteq\fun{txMD}~\var{tx}
      \\
      (\var{mdh}=\Nothing \land \var{md}=\Nothing)
      \lor
      (\var{mdh}=\fun{hashMD}~\var{md})
      \\~\\
      {
        \begin{array}{r}
          \var{slot}\\
          \var{pp}\\
          \var{stkCreds}\\
          \var{stpools}\\
          \var{genDelegs}\\
        \end{array}
      }
      \vdash \var{utxoSt} \trans{\hyperref[fig:rules:utxo-shelley]{utxo}}{tx}
      \var{utxoSt'}\\
    }
    {
      \begin{array}{r}
        \var{slot}\\
        \var{pp}\\
        \var{stkCreds}\\
        \var{stpools}\\
        \var{genDelegs}\\
      \end{array}
      \vdash \var{utxoSt} \trans{utxow}{tx} \varUpdate{\var{utxoSt'}}
    }
  \end{equation}
  \caption{UTxO with witnesses inference rules for GoguenTx}
  \label{fig:rules:utxow-goguen}
\end{figure}

The second rule applies when we have $tx \in \ShelleyTx$, see
Figure~\ref{fig:rules:utxow-goguen}. In this case, we explicitly use the Shelley
era functions and types, exactly as defined in the Shelley spec, to perform all the
witnessing checks in the Shelley UTXOW rule. We must adhere to the Shelley
witnessing rules on the transaction as is because the VK or multisig
script signatures will not validate if we transform the transaction in any way.

Having performed the witnessing checks on the original transaction, we use
the $\fun{toGoguenTx}$ function (defined in Figure
~\ref{fig:functions:to-shelley}) to strip all of the witnessing data off the
transaction, and convert it to a Goguen type transaction. We keep the metadata and
put default trivial values in any fields that are added to the transaction type
of the Goguen era. We use this transformed version as the signal for the
UTXO transition (which again determines the state change of this UTXOW rule).

\begin{figure}
  \begin{equation}
    \label{eq:utxo-witness-inductive-shelley}
    \inference[UTxO-witS]
    {
      \var{txb}\leteq\txbody{tx} &
      \var{txw}\leteq\fun{txwits}~{tx} &
       \var{tx}~\in~\ShelleyTx \\~\\
      (utxo, \wcard, \wcard, \wcard) \leteq \var{utxoSt} \\
      \var{witsKeyHashes} \leteq \{\fun{hashKey}~\var{vk} \vert \var{vk} \in
      \dom (\txwitsVKey{txw}) \}\\~\\
      \forall \var{hs} \mapsto \var{validator} \in \fun{txwitsScript}~{txw},\\
      \fun{hashScript}~\var{validator} = \var{hs} \wedge
      \fun{validateScript}~\var{validator}~\var{tx}\\~\\
      \fun{scriptsNeeded}~\var{utxo}~\var{tx} = \dom (\fun{txwitsScript}~{txw})
      \\~\\
      \forall \var{vk} \mapsto \sigma \in \txwitsVKey{txw},
      \mathcal{V}_{\var{vk}}{\serialised{\txbody{tx}}}_{\sigma} \\
      \fun{witsVKeyNeeded}~{utxo}~{tx}~{genDelegs} \subseteq witsKeyHashes
      \\~\\
      genSig \leteq
      \left\{
        \fun{hashKey}~gkey \vert gkey \in\dom{genDelegs}
      \right\}
      \cap
      \var{witsKeyHashes}
      \\
      \left\{
        c\in\txcerts{txb}~\cap\DCertMir
      \right\} \neq\emptyset \implies \vert genSig\vert \geq \Quorum \wedge
      \fun{d}~\var{pp} > 0
      \\~\\
      \var{mdh}\leteq\fun{txMDhash}~\var{txb}
      &
      \var{md}\leteq\fun{txMD}~\var{tx}
      \\
      (\var{mdh}=\Nothing \land \var{md}=\Nothing)
      \lor
      (\var{mdh}=\fun{hashMD}~\var{md})
      \\~\\
      \var{txg}~\leteq~\fun{toGoguenTx}~{tx} \\~\\
      {
        \begin{array}{r}
          \var{slot}\\
          \var{pp}\\
          \var{stkCreds}\\
          \var{stpools}\\
          \var{genDelegs}\\
        \end{array}
      }
      \vdash \var{utxoSt} \trans{\hyperref[fig:rules:utxo-shelley]{utxo}}{txg}
      \var{utxoSt'}\\
    }
    {
      \begin{array}{r}
        \var{slot}\\
        \var{pp}\\
        \var{stkCreds}\\
        \var{stpools}\\
        \var{genDelegs}\\
      \end{array}
      \vdash \var{utxoSt} \trans{utxow}{tx} \varUpdate{\var{utxoSt'}}
    }
  \end{equation}
  \caption{UTxO with witnesses inference rules for ShelleyTx}
  \label{fig:rules:utxow-shelley}
\end{figure}

\clearpage
