\hypersetup{
  pdftitle={A Formal Specification of the Cardano Ledger},
  breaklinks=true,
  bookmarks=true,
  colorlinks=false,
  linkcolor={blue},
  citecolor={blue},
  urlcolor={blue},
  linkbordercolor={white},
  citebordercolor={white},
  urlbordercolor={white}
}

  \cleardoublepage%
  \tableofcontents%

  \listoffigures%
  \clearpage%

  \begin{changelog}
        \change{2019/10/08}{Jared Corduan, Polina Vinogradova and Matthias G\"udemann}{FM (IOHK)}{Initial version (0.1).}
        \change{2019/10/08}{Kevin Hammond}{FM (IOHK)}{Added cover page.}
        \change{2020/11/17}{Jared Corduan}{FM (IOHK)}
          {Removed unused deliverable outline image,
           set version to 1.0,
           and changed the reward calculation so that $\eta$
             takes $d$ into account.}
        \change{2021/05/17}{Jared Corduan}{FM (IOHK)}{Added Example Illustration.}
        \change{2021/06/14}{Jared Corduan}{FM (IOHK)}
          {Added an errata section,
           fixed a typo in leader check and in the LEDGERS rule index,
           and synced the reward calculation with the implementation}
        \change{2021/06/17}{Jared Corduan}{FM (IOHK)}
          {Allow the pool influence parameter $a_0$ to be zero,
          remove all mentions of deposit decay, sync varibale names with code.}
        \change{2021/08/27}{Jared Corduan}{FM (IOHK)}{Fixed definitions in the header-only validation properties.
          CHAIN transition did not need to use previous protocol parameters.}
        \change{2021/10/08}{Jared Corduan}{FM (IOHK)}{The function createRUpd
          should get the pool parameters from the go snapshot.
          The TICKN rule was missing from the dependency diagram.}
        \change{2021/11/08}{Jared Corduan}{FM (IOHK)}{Fixed typo in the description of variable length encodings.}
      \end{changelog}
      \clearpage%
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}

\title{A Formal Specification of the Cardano Ledger}

\author{Jared Corduan  \\ {\small \texttt{jared.corduan@iohk.io}} \\
   \and Polina Vinogradova \\ {\small \texttt{polina.vinogradova@iohk.io}} \\
   \and Matthias G\"udemann  \\ {\small \texttt{matthias.gudemann@iohk.io}}}

%\date{}

\maketitle

\begin{abstract}
This document provides a formal specification of the Cardano ledger for use in the upcoming Shelley implementation.
It is intended to underpin a Haskell executable specification that will be the basis of the initial
Shelley release, and represents a core design and quality assurance document.
It will be used to define properties and tests, and to provide the basis for strong formal assurance
using mathematical proof techniques.
The document defines the rules for extending the ledger with transactions
that will affect both UTxO and stake delegation.
Key properties that have been identified include the preservation of balances, absence of double spend, stakepool registration,
and reward splitting.
\end{abstract}

\section*{List of Contributors}
\label{acknowledgements}

Nicol\'as Arqueros,
Dan Bornside,
Nicholas Clarke,
Duncan Coutts,
Ruslan Dudin,
Sebastien Guillemot,
Kevin Hammond,
Vincent Hanquez,
Ru Horlick,
Michael Hueschen,
Christian Lindgren,
Yun Lu,
Philipp Kant,
Jean-Christophe Mincke,
Damian Nadales,
Ashish Prajapati,
Nicolas Di Prima,
Andrew Westberg.
