\section{Output Size}
\label{sec:value-size}

In Shelley, the protocol parameter $\var{minUTxOValue}$ is used to
disincentivize attacking nodes by putting many small outputs on the
UTxO, permanently blocking memory. Because members of the $\Value$
type can be arbitrary large, the ledger requires a $\UTxO$ entry to
contain a minimum amount of Ada, proportional to its size.

There is also another $\Value$ size consideration with respect to spendability
of an output. The restriction on the total serialized size of the transaction (set
by the parameter $\var{maxTxSize}$) serves as an implicit upper bound on the
size of a $\Value$ contained in an output of a transaction. Without tighter
limits on the output $\Value$ size, one of the following situations could arise,
causing the output to be come unspendable (these are just a two examples) :

\begin{itemize}
  \item The script locking the very large $\Value$-containing UTxO is too large
  to fit inside the transaction alongside the $\Value$ itself while still respecting
  the max transaction size
  \item The large $\Value$ cannot be split into several outputs, because the
  outputs are impossible to fit inside a single transaction
\end{itemize}

Figure \ref{fig:size-helper} contains abstract and helper functions
used in calculating the in-memory and serialized representation
sizes of $\Value$ terms.

\begin{figure*}[h]
  \emph{Abstract Functions}
  %
  \begin{align*}
    & \fun{serSize} \in \Value \to \MemoryEstimate \\
    & \text{Gives the size of the serialized representation of a $\Value$}
    \nextdef
    & \fun{anameLen} \in \AssetName \to \MemoryEstimate \\
    & \text{Returns the length (in bytes) of an asset name}
  \end{align*}
  %
  \emph{Helper Functions}
  \begin{align*}
    & \fun{numAssets} \in \Value \to \N \\
    & \fun{numAssets}~{vl}~=~\|~\{~(pid, an)~\vert~pid~(\mapsto~(an~\mapsto~\wcard))~\in~vl~\}~\| \\
    & \text{Returns the number of distinct asset IDs in a $\Value$}
    & \nextdef
    & \fun{sumALs} \in \Value \to \N \\
    & \fun{sumALs}~{vl}~=~ \sum_{\{~an~\vert~\wcard~(\mapsto~(an~\mapsto~\wcard))~\in~vl~\}} \fun{anameLen}~an \\
    & \text{Returns the sum of the lengths (in bytes) of distinct asset names in a $\Value$}
    & \nextdef
    & \fun{numPids} \in \Value \to \N \\
    & \fun{numPids}~{vl} ~=~ \|~\supp~{vl}~\| \\
    & \text{The number of policy IDs in a $\Value$}
  \end{align*}
  \caption{Value Size}
  \label{fig:size-helper}
\end{figure*}

The function $\fun{serSize}$, unlike the functions $\fun{size}$ and $\fun{utxoEntrySize}$
    (explained below), is left abstract. It returns the actual number of bytes a $\Value$
    occupies in the serialized transaction (and hence is implementation-dependent).

    One reason for this design decision is that the $\fun{serSize}$ function is used constrain
    the serialized representation of the transaction (in particular, the size
    of $\Value$ terms in outputs), whereas the min-Ada requirement is about
    the in-memory representation size. Moreover, a transparently-calculated size estimate
    is not necessary for limiting the size of values in outputs, since this size-bound
    check does not place any additional accounting/monetary constranits on transaction construction,
    unlike the min-Ada requirement.

Figure \ref{fig:min-val-calc} gives the types of constants used in the estimation
of the size of a UTxO entry, and the associated min-Ada-value.

\begin{figure*}[h]
  \emph{Constants}
  \begin{equation*}
    \begin{array}{lcl}
      (k_0, k_1, k_2, k_3, k_4) & \in & \N \times \N \times \N \times \N \times \N \\
      \mathsf{UtxoEntrySizeWithoutVal} & \in & \MemoryEstimate \\
      \mathsf{AdaOnlyUTxOSize} & \in & \MemoryEstimate \\
      \mathsf{MaxValSize} & \in & \MemoryEstimate \\
    \end{array}
  \end{equation*}
  %
  \emph{Size and Min-Ada Functions}
  \begin{align*}
    & \fun{size} \in \Value \to \MemoryEstimate \\
    & \fun{size}~\var{vl} ~=~
    \begin{cases}
      k_0 & \fun{adaOnly}~vl~=~vl \\
      k_1 + \lfloor~ (((\fun{numAssets}~vl) * k_2) + (\fun{sumALs}~vl) & \\
      ~~~~~~ + ((\fun{numPids}~vl) * k_3) + (k_4 - 1)))~ /~ k_4~\rfloor & \text{otherwise} \\
    \end{cases} \\
    & \text{Calculate the size of a $\Value$}
    \nextdef
    & \fun{adaPerUTxOWord}~\in \PParams \to \Coin \\
    & \fun{adaPerUTxOWord}~pp = \lfloor~ (\fun{minUTxOValue}~pp~/~ \mathsf{adaOnlyUTxOSize})~ \rfloor \\
    & \text{Calculate the cost of storing a memory unit of data as a UTxO entry}
    \nextdef
    & \fun{utxoEntrySize} \in \TxOut \to \MemoryEstimate \\
    & \fun{utxoEntrySize}~\var{out} = \mathsf{utxoEntrySizeWithoutVal} + \fun{size}~ (\fun{getValue}~out) \\
    & \text{Calculate the size of a UTxO entry}
\end{align*}
\caption{Value Size}
\label{fig:min-val-calc}
\end{figure*}

The $\fun{size}$ function returns the estimated size of a $\Value$ term. It is constant in the
case where the value contains only Ada. If there are other types of tokens in the
term, the size depends on

\begin{itemize}
  \item the number of distinct asset types (asset IDs)
  \item the number of distinct policy IDs, and
  \item the sum of the lengths of distinct asset names of the tokens.
\end{itemize}

The parameter $\fun{minUTxOValue}$ specifies the min-Ada value for a UTxO containing
only Ada. This type of UTxO varies in size somewhat (eg. Byron style addresses
may be a different length than Shelley ones), but we estimate the size of the most commonly
used type of Ada-only UTxO as the constant value $\mathsf{adaOnlyUTxOSize}$.
This constant is in fact an upper bound on UTxOs which have have only Shelley credentials.

We use this size estimate
to calculate what $\fun{minUTxOValue}$ implies to be the min-Ada value requirement
\emph{per word} of UTxO data.

The function $\fun{adaPerUTxOWord}$ performs this calculation by dividing the
min-Ada value by the Ada-only UTxO size.

The $\mathsf{utxoEntrySizeWithoutVal}$ is the constant representing
the size of a UTxO entry, not counting the size of the $\Value$ term it contains.
Here, again, the actual size of a UTxO (excluding the $\Value$) can vary, but
we use an upper bound on the size of Shelley-credential UTxOs.

The function $\fun{utxoEntrySize}$ estimates the size of an arbitrary ShelleyMA-era
UTxO. It adds the size estimate of the $\Value$ term in a UTxO and the
$\mathsf{utxoEntrySizeWithoutVal}$ constant.

The constants used in the implementation of the Alonzo era are as follows :

\begin{itemize}
  \item $(k_0, k_1, k_2, k_3, k_4) = (2, 6, 12, 28, 8)$
  \item $\mathsf{utxoEntrySizeWithoutVal} = 27$ words (8 bytes)
  \item $\mathsf{adaOnlyUTxOSize} = 27$ words (8 bytes)
  \item $\mathsf{MaxValSize} = 4000$ bytes, ie. 500 words.
\end{itemize}
