\lstset{ frame=tb,
       , language=Haskell
       , basicstyle=\footnotesize\ttfamily,
       , keywordstyle=\color{blue!80},
       , commentstyle=\itshape\color{purple!40!black},
       , identifierstyle=\bfseries\color{green!40!black},
       , stringstyle=\color{orange},
       }

\lstMakeShortInline[columns=fixed]`

\subsection{Aspects that we need to model}
\label{sec:aspects-to-model}

\begin{description}
\item[Authentication] Update proposals and votes are authenticated (properly
  signed).
\item[Authorization] Only genesis keys (via deleg certs) can post update
  proposals.
  \begin{itemize}
  \item Only then they can vote on them.
  \end{itemize}
\item[Voting deadlines] voting end when a majority of the voters (4/7 if we
assume no stake) agree on the proposal.
\item[Block versions] (= protocol versions)
\item[Soft-forks] a protocol version changes according to the fork-resolution
  rule (75\% of stake create blocks with new-version).
\item[{Hard-forks}] ??? Do we need to model anything here?.
\end{description}

\subsection{Information in the ledger state}
\label{sec:information-in-ledger-state}

The ledger state has to expose some parameters of the protocol version to its
clients. In the current implementation these parameters are kept in the
`BlockVersionData` structure.

\begin{lstlisting}
  data BlockVersionData = BlockVersionData { ... }
\end{lstlisting}

The following parameters are likely to be needed by the consumers of the ledger
layer:

\begin{itemize}
\item Slot duration (`bvdSlotDuration`)
\item Maximum block size (`bvdMaxBlockSize`)
\item Maximum header size (`bvdMaxHeaderSize`)
\item Maximum transaction size (`bvdMaxTxSize`)
\item Maximum proposal size (`bvdMaxProposalSize`)
\item Transaction fee policy (`bvdTxFeePolicy`)
\end{itemize}

The following parameters will be assumed to be constant:
\begin{itemize}
\item MPC threshold (`bvdMpcThd`)
\item Heavy delegation threshold (`bvdHeavyDelThd`)
\item Update vote threshold (`bvdUpdateVoteThd`)
\item Update proposal threshold (`bvdUpdateProposalThd`)
\item Soft fork rule parameters (`bvdSoftforkRule`)
\end{itemize}

At the moment we don't know whether we need these:

\begin{itemize}
\item Script Version (`bvdScriptVersion`)
\item Update implicit (`bvdUpdateImplicit`)
\end{itemize}

Finally, the `bvdUnlockStakeEpoch` field of `BlockVersionData` does not need to
be modeled.
